---
title: "MethScope-Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MethScope-Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(device = "png")
knitr::opts_chunk$set(
  fig.ext = "png",
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MethScope)
```

## Overview

This tutorial demonstrates a standard MethScope workflow:

- Generate MRMP embeddings from `.cg` input
- Perform cell-type annotation with a pre-trained model
- Train a custom model
- Visualize prediction quality
- Run deconvolution and unsupervised clustering

Data and preprocessing resources:

- Example/reference data: <https://github.com/zhou-lab/methscope_data>
- Generate `.cg` input files: <https://zhou-lab.github.io/YAME/>
- Interpret MRMP patterns: <https://www.bioconductor.org/packages/release/bioc/html/knowYourCG.html>

## 1) Generate cell/pixel/sample MRMP embeddings

Use `GenerateInput()` to transform a query `.cg` file into a sample-by-MRMP matrix.

```{r, eval=FALSE}
# Paths to your .cg and .cm files
example_file <- "example.cg"
reference_pattern <- "Liu2021_MouseBrain.cm"
input_pattern <- GenerateInput(example_file, reference_pattern)
```

For large datasets, split `.cg` files and process in parallel:
<https://zhou-lab.github.io/YAME/docs/subset.html>

## 2) Perform cell-type annotation

Use a built-in pre-trained model for fast annotation.
Available built-in examples include:

- `MethScope:::Liu2021_MouseBrain_P1000`
- `MethScope:::Zhou2025_HumanAtlas_P1000`

```{r, eval=FALSE}
prediction_result <- PredictCellType(
  MethScope:::Liu2021_MouseBrain_P1000,
  input_pattern
)
```

## 3) Train a custom classification model

Train your own XGBoost model using known labels.
`cell_type_label` must align with rows of `input_pattern`.

```{r, eval=FALSE}
trained_model <- Input_training(
  input_pattern,
  cell_type_label
)
```

You can also tune hyperparameters automatically:

```{r, eval=FALSE}
trained_model_cv <- Input_training(
  input_pattern,
  cell_type_label,
  cross_validation = TRUE
)
```

## 4) Visualize prediction results

```{r, eval=FALSE}
umap_plot <- PlotUMAP(input_pattern, prediction_result)
# cell_type_label is the true label vector
PlotConfusion(prediction_result, cell_type_label)
PlotF1(prediction_result, cell_type_label)
```

## 5) Cell-type deconvolution

Estimate cell-type proportions using `nnls_deconv()`.
Reference matrices are available in:
<https://github.com/zhou-lab/methscope_data>

```{r, eval=FALSE}
reference_pattern <- "Liu2021_MouseBrain.cm"
reference_input <- readRDS("2021Liu_reference_pattern.rds")
cell_proportion <- nnls_deconv(reference_input, input_pattern)
```

## 6) Unsupervised clustering (Seurat example)

After generating `input_pattern`, you can cluster with standard single-cell pipelines.

```{r, eval=FALSE}
Pattern.obj <- CreateSeuratObject(counts = t(input_pattern), assay = "DNAm")
VariableFeatures(Pattern.obj) <- rownames(Pattern.obj[['DNAm']])
DefaultAssay(Pattern.obj) <- "DNAm"
Pattern.obj <- NormalizeData(Pattern.obj, assay = "DNAm", verbose = FALSE)
Pattern.obj <- ScaleData(Pattern.obj, assay = "DNAm", verbose = FALSE)
# You can also directly use the initial counts matrix
Pattern.obj@assays$DNAm@layers$scale.data <- as.matrix(Pattern.obj@assays$DNAm@layers$counts)
Pattern.obj <- RunPCA(Pattern.obj, assay = "DNAm", reduction.name = "mpca", verbose = FALSE)
Pattern.obj <- FindNeighbors(Pattern.obj, reduction = "mpca", dims = 1:30)
Pattern.obj <- FindClusters(Pattern.obj, verbose = FALSE, resolution = 0.7)
Pattern.obj <- RunUMAP(Pattern.obj, reduction = "mpca", reduction.name = "meth.umap", dims = 1:30)
```

## Related resources

- MRMP reference tutorial: <https://zhou-lab.github.io/MethScope/articles/MethScope-MRMP.html>
- Full package site: <https://zhou-lab.github.io/MethScope/>
