---
title: "MethScope-Input"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MethScope-Input}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

MethScope uses YAME `.cg` files as methylation input. A `.cg` file is a compact
binary representation of methylation data in a fixed CpG order. Before running
`GenerateInput()`, the query `.cg` file and the MRMP reference `.cm` file must be
based on the same genome build and CpG coordinate order.

This tutorial shows common input-preparation patterns for:

- BED-like methylation files
- ALLC files
- beta or methylation fraction tables
- binary presence/absence tracks

For complete YAME command documentation, see <https://zhou-lab.github.io/YAME/>.
YAME's storage guide describes the supported CX formats and the requirement that
input rows match the reference CpG coordinates exactly:
<https://zhou-lab.github.io/YAME/docs/storage.html>.

For ALLC files, this tutorial follows the ALLCools definition:
<https://lhqing.github.io/ALLCools/start/input_files.html>.

## Install command-line dependencies

Install YAME and bedtools before converting files.

```{bash, eval=FALSE}
conda install -c bioconda yame bedtools
```

## Choose the correct CpG reference

Use the CpG reference that matches the genome build of your data and the MethScope
MRMP reference. For example, use an `mm10` CpG reference with
`mm10_Liu2021.cm`, and use an `hg38` CpG reference with `hg38_Zhou2025.cm` or
`hg38_Loyfer2023.cm`.

The examples below assume that you have a YAME CpG reference file called
`cpg_reference.cr`. Convert it to BED once:

```{bash, eval=FALSE}
yame unpack cpg_reference.cr | gzip > cpg_reference.bed.gz
```

The BED file created from the `.cr` reference defines the required row order for
the output `.cg` file.

## Convert M/U count BED files

Use YAME format 3 (`-f3`) when the input has methylated and unmethylated read
counts. This is the recommended format for bisulfite sequencing data when read
depth is available.

Expected input:

```text
chr1    3000826    3000827    4    6
chr1    3001006    3001007    0    8
```

where columns are:

```text
chrom    start    end    M    U
```

Align the file to the reference CpG coordinates and pack:

```{bash, eval=FALSE}
bedtools sort -i sample_mu.bed > sample_mu.sorted.bed

bedtools intersect \
  -a cpg_reference.bed.gz \
  -b sample_mu.sorted.bed \
  -loj \
  -sorted | \
awk 'BEGIN{OFS="\t"} {if ($8==".") print 0,0; else print $8,$9}' | \
yame pack -f3 - > sample.cg
```

Check the result:

```{bash, eval=FALSE}
yame info sample.cg
yame summary sample.cg
```

## Convert ALLC files

ALLC files are tab-separated base-resolution cytosine count tables. In the
ALLCools convention, each row has 7 mandatory columns and no header:

```text
chromosome    position    strand    sequence_context    mc    cov    methylated
```

where `position` is 1-based, `mc` is the count of reads supporting methylation,
and `cov` is total read coverage. For MethScope, convert CG-context rows to
YAME format 3 by setting:

```text
M = mc
U = cov - mc
```

The command below keeps rows whose sequence context begins with `CG`, converts
the 1-based ALLC position to a 0-based BED interval, derives `U`, aligns the
result to the CpG reference, and packs the final two-column `M U` stream.

```{bash, eval=FALSE}
zcat sample.allc.tsv.gz | \
awk 'BEGIN{OFS="\t"}
     $4 ~ /^CG/ {
       chrom = $1
       start = $2 - 1
       end = $2
       m = $5
       u = $6 - $5
       if (u < 0) u = 0
       print chrom, start, end, m, u
     }' | \
bedtools sort -i - > sample_mu.sorted.bed

bedtools intersect \
  -a cpg_reference.bed.gz \
  -b sample_mu.sorted.bed \
  -loj \
  -sorted | \
awk 'BEGIN{OFS="\t"} {if ($8==".") print 0,0; else print $8,$9}' | \
yame pack -f3 - > sample.cg
```

If your ALLC file is not compressed, replace `zcat sample.allc.tsv.gz` with
`cat sample.allc.tsv`.

## Convert beta or methylation fraction files

Use YAME format 4 (`-f4`) when the input already contains methylation fractions
or beta values. Missing CpGs should be encoded as `NA`.

Expected input:

```text
chr1    3000826    3000827    0.42
chr1    3001006    3001007    0.81
```

Convert to `.cg`:

```{bash, eval=FALSE}
bedtools sort -i sample_beta.bed > sample_beta.sorted.bed

bedtools intersect \
  -a cpg_reference.bed.gz \
  -b sample_beta.sorted.bed \
  -loj \
  -sorted | \
awk 'BEGIN{OFS="\t"} {if ($8==".") print "NA"; else print $8}' | \
yame pack -f4 - > sample_beta.cg
```

Format 4 stores methylation fractions but does not retain read depth.

## Convert binary BED tracks

Use YAME format 0 (`-fb`) when each CpG is represented as present or absent.
This is useful for binary tracks, peak overlaps, or binarized methylation states.

```{bash, eval=FALSE}
bedtools sort -i binary_track.bed > binary_track.sorted.bed

bedtools intersect \
  -a cpg_reference.bed.gz \
  -b binary_track.sorted.bed \
  -sorted \
  -c | \
cut -f4 | \
awk '{if ($1 > 0) print 1; else print 0}' | \
yame pack -fb - > binary_track.cg
```

## Use the converted file in MethScope

After creating `sample.cg`, use an MRMP reference `.cm` from the same genome
build.

```{r, eval=FALSE}
library(MethScope)

query_file <- "sample.cg"
reference_pattern <- "mm10_Liu2021.cm"

input_pattern <- GenerateInput(query_file, reference_pattern)

model <- Liu2021_MouseBrain_P1000()
prediction_result <- PredictCellType(model, input_pattern)
```

For a full MethScope test, clone the GitHub repository and run the tutorial with
`inst/extdata/example.cg` and `inst/extdata/mm10_Liu2021.cm`.

## Troubleshooting

If `GenerateInput()` fails or returns unexpected values, check these items first:

- The query `.cg` and reference `.cm` use the same genome build.
- The query `.cg` was packed against the same CpG coordinate order used by the
  `.cm` reference.
- `yame info sample.cg` reports the expected number of CpGs.
- `yame summary sample.cg` shows nonzero coverage for real sequencing data.
- For format 3 input, no-coverage CpGs should be `0 0`.
- For format 4 input, missing values should be `NA`.

## Related resources

- YAME documentation: <https://zhou-lab.github.io/YAME/>
- YAME storage and format guide: <https://zhou-lab.github.io/YAME/docs/storage.html>
- ALLCools input-file documentation: <https://lhqing.github.io/ALLCools/start/input_files.html>
- MethScope tutorial: <https://zhou-lab.github.io/MethScope/articles/MethScope-Tutorial.html>
