---
title: "20250404_spatial"
output: html_document
date: "2025-04-04"
---

```{r}
library(Seurat)
library(ArchR)
reference_pattern <- system.file("extdata", "Liu2021_MouseBrain.cm", package = "MethScope")
input_pattern <- GenerateInput("/home/fuh1/zhou_lab/projects/20230531_Spatial_methseq/pileup/SpMETSLE14DM/SpMETSLE14DM.cg",reference_pattern)
#input_pattern <- GenerateInput("/home/fuh1/zhou_lab/projects/20230531_Spatial_methseq/pileup/SpMETSLB14DM/SpMETSLB14DM.cg",reference_pattern)
valid_barcodes <- readRDS("/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20240331_E14D_barcodesK.rds")
#valid_barcodes <- readRDS("/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20240331_B14D_barcodesK.rds")
input_pattern <- input_pattern[rownames(input_pattern) %in% valid_barcodes,]
#saveRDS(input_pattern,"~/cell_type/20250408_P21_input.rds")

reference_input <- GenerateInput("/mnt/isilon/zhou_lab/projects/20230727_all_public_WGBS/mm10/2021_LiuMajorTypePseudo.cg",reference_pattern)

input_pattern_imputed <- imputeRowMean(input_pattern,na_percent = 30)
reference_input_imputed <- imputeRowMean(reference_input,na_percent = 30)
cell_proportion <- nnls_deconv(reference_input_imputed,input_pattern_imputed)

# top_celltypes <- apply(cell_proportion, 2, function(x) {
#   rownames(cell_proportion)[which.max(x)]
# })
# 
# spatial_plot(as.factor(top_celltypes),colnames(cell_proportion))

# training_model <- readRDS("/home/fuh1/intermediate/20250212_Liu2021_training_model.rds")
# prediction_result <- PredictCellType(training_model,input_pattern,smooth = F)  

input_pattern_imputed <- imputeRowMean(input_pattern)
#input_pattern_imputed <- cbind(input_pattern_imputed,prediction_result[,1:length(training_model$cell_type)])
input_pattern_imputed <- cbind(input_pattern_imputed,t(cell_proportion))

P21_brain <- CreateSeuratObject(counts = t(input_pattern_imputed), assay = "DNAm")
VariableFeatures(P21_brain) <- rownames(P21_brain[['DNAm']])
DefaultAssay(P21_brain) <- "DNAm"
P21_brain <- NormalizeData(P21_brain, assay = "DNAm", verbose = FALSE)
P21_brain <- ScaleData(P21_brain, assay = "DNAm", verbose = FALSE)
#P21_brain@assays$DNAm@layers$scale.data <- as.matrix(P21_brain@assays$DNAm@layers$counts)

P21_brain <- RunPCA(P21_brain,assay="DNAm",reduction.name = 'mpca', verbose = FALSE)
P21_brain <- FindNeighbors(P21_brain, reduction = "mpca", dims = 1:30)
P21_brain <- FindClusters(P21_brain, verbose = FALSE, resolution = 0.7)
P21_brain <- RunUMAP(P21_brain, reduction = "mpca",  reduction.name = "meth.umap", dims = 1:30)

custom_colors <- c("#8A9FD1","#FEE500","#89C75F","#D51F26","#F47D2B","#D8A767","#89288F","#208A42")
P21_brain$DNAm_snn_res.0.7 <- paste0("D",P21_brain$DNAm_snn_res.0.7)
DimPlot(P21_brain, reduction = 'meth.umap', repel = TRUE, group.by = 'DNAm_snn_res.0.7',cols=custom_colors, pt.size = 0.7,alpha = 0.7)  + NoGrid() + ggtitle("")
#ggsave("~/figures/fuh1/2025707_P21_MethScope_umap.pdf",width=3,height=3)

### E14D
custom_colors <- c("#8A9FD1","#89288F","#FEE500","#F47D2B","#D51F26","#208A42")
DimPlot(P21_brain, reduction = 'meth.umap', repel = TRUE, group.by = 'DNAm_snn_res.0.7', cols=custom_colors,pt.size = 0.7,alpha = 0.7)  + NoGrid() + ggtitle("")
ggsave("~/figures/fuh1/20250616_E14D_MethScope_umap.pdf",width=3,height=3)

partition <- Idents(P21_brain)
#partition <- P21_brain$MethScope

spatial_barcode <- colnames(P21_brain)
spatial_plot(partition,spatial_barcode) + scale_color_manual(values=custom_colors,na.value="white",na.translate = FALSE,labels = c("D0","D1","D2","D3","D4","D5")) + labs(color ="")
#ggsave("~/figures/fuh1/20250616_E14D_MethScope.pdf",width=4.5,height=4)
#saveRDS(P21_brain,"~/cell_type/20250408_P21_input_obj.rds")


custom_colors <- c("#FEE500","#8A9FD1","#D8A767","#89288F","#D51F26")
P21_brain_orig <- readRDS("~/tmp/20231115_spatial_window/spatial.obj_06022024.rds")
DefaultAssay(P21_brain_orig) <- 'DNAm_frac_CG'
DimPlot(P21_brain_orig, reduction = 'meth.umap', repel = TRUE, group.by = 'DNAm_CG_snn_res.0.6',cols=custom_colors,pt.size = 0.7,alpha = 0.7) + NoGrid() + ggtitle("")
#ggsave("~/figures/fuh1/20250213_P21_VMR_umap.pdf",width=3,height=3)

P21_brain_orig_barcode <- sub("\\..*$", "", colnames(P21_brain_orig))
P21_brain_orig_barcode <- paste0(substr(P21_brain_orig_barcode,9,16), substr(P21_brain_orig_barcode,1,8))

digits <- as.integer(gsub("D", "", P21_brain_orig@meta.data$DNAm_CG_snn_res.0.6))

spatial_plot(as.factor(digits),valid_barcodes) + scale_color_manual(values=custom_colors,na.value="white",na.translate = FALSE)
#ggsave("~/figures/fuh1/20250213_P21_VMR.pdf",width=4.5,height=4)

cell_type <- "CT-L6"
spatial_plot(cell_proportion[cell_type,],names(cell_proportion[cell_type,]))+labs(color = cell_type)
ggsave(paste0("~/figures/fuh1/20250501_P21_",cell_type,".pdf"),width=4.5,height=4)

```

### plot cell type proportions
```{r}
prop_df <- as.data.frame(t(cell_proportion))
prop_df$group <- Idents(P21_brain)  # add group label

avg_props <- prop_df %>%
  group_by(group) %>%
  dplyr::summarise(across(.cols = everything(), .fns = mean))

avg_props_mat <- as.data.frame(t(avg_props[,-1]))  # drop group column before transposing
colnames(avg_props_mat) <- avg_props$group
rownames(avg_props_mat) <- rownames(cell_proportion)

avg_props_mat_filtered <- avg_props_mat[rowSums(avg_props_mat[,-ncol(avg_props_mat)]) > 0.1, ]
avg_props_norm <- sweep(avg_props_mat_filtered, 2, colSums(avg_props_mat_filtered), FUN = "/")


avg_props_norm$CellType <- rownames(avg_props_norm)
avg_props_long <- melt(avg_props_norm, id.vars = "CellType", variable.name = "Group", value.name = "Proportion")


#cols <- DiscretePalette(n_clusters, palette = "alphabet", shuffle = FALSE)

n_clusters <- length(unique(avg_props_long$CellType))
cols <- as.character(ArchRPalettes$grove[seq_len(n_clusters)])

ggplot(avg_props_long, aes(x = Group, y = Proportion, fill = CellType)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ylab("Average Proportion") +
  xlab("")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.ticks.x = element_blank(), axis.ticks.y = element_blank(),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(title = "")) + scale_fill_manual(values=cols)
ggsave("~/figures/fuh1/20250213_P21_MethScope_proportion.pdf",width=5,height=4)


# heatmap

library(pheatmap)
cell_proportion_z <- t(scale(t(cell_proportion), center = TRUE, scale = TRUE))
cell_proportion_z_clean <- cell_proportion_z[!apply(cell_proportion_z, 1, function(x) all(is.na(x))), ]
cell_proportion_z_clean <- cell_proportion_z_clean[apply(cell_proportion_z_clean, 1, function(x) all(abs(x) <= 20)), ]
cell_proportion_z_clean <- pmax(pmin(cell_proportion_z_clean, 3), -3)
prop_mat <- cell_proportion_z_clean

cell_types <- c("ASC","DG-po","CT-L6","IT-L23","DG","ODC","CA1","CA3","MGC","EC","OLF-Exc","PAL-Inh")
prop_mat <- prop_mat[rownames(prop_mat) %in% cell_types,]

# Reorder samples by cluster
cluster_vec <- Idents(P21_brain)
cluster_order <- order(cluster_vec)
prop_ordered <- prop_mat[, cluster_order]  # reorder columns
clusters <- cluster_vec[cluster_order]

# Optional: assign a color per cluster
cluster_levels <- unique(clusters)
cluster_colors <- setNames(c("#8A9FD1","#FEE500","#89C75F","#D51F26","#F47D2B","#D8A767","#89288F","#208A42"), cluster_levels)

# Prepare annotation for columns (samples)
annotation_col <- data.frame(Cluster = clusters)
rownames(annotation_col) <- colnames(prop_ordered)

# Plot the heatmap
library(pheatmap)
pdf("~/figures/fuh1/20250707_P21_proportion.pdf",width=6,height=4.5)
pheatmap(prop_ordered,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         annotation_col = annotation_col,
         annotation_colors = list(Cluster = cluster_colors),
         show_colnames = FALSE,
         fontsize_row = 10,
         scale = "none",
         main = "",
         color = hcl.colors(50, "Viridis"))
dev.off()

### cluster columns 
# Prepare annotation for columns
cluster_levels <- unique(clusters)
cluster_colors <- setNames(
  c("#8A9FD1", "#FEE500", "#89C75F", "#D51F26", "#F47D2B", "#D8A767", "#89288F", "#208A42"),
  cluster_levels
)

# Annotation dataframe for columns
annotation_col <- data.frame(Cluster = clusters)
rownames(annotation_col) <- colnames(prop_ordered)

# Define annotation with colors
col_ha <- HeatmapAnnotation(
  df = annotation_col,
  col = list(Cluster = cluster_colors),
  show_annotation_name = TRUE
)

library(ComplexHeatmap)
min_val <- min(prop_ordered)
max_val <- max(prop_ordered)
# Get all cluster levels you want, in order
desired_order <- 0:7
clusters_named <- factor(
  clusters,
  levels = desired_order,
  labels = paste0("Cluster ", desired_order)
)

pdf("~/figures/fuh1/20250707_P21_proportion.pdf",width=6,height=4.5)
Heatmap(
  as.matrix(prop_ordered),
  name = "Proportion",
  top_annotation = col_ha,
  cluster_rows = FALSE,
  cluster_columns = TRUE,         # This will cluster within each cluster
  column_split = clusters_named,        # Split by cluster group
  column_title = NULL,   
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 10),
  col = colorRamp2(
  seq(min_val, max_val, length.out = 50),
  hcl.colors(50, "Inferno")
  ),
  show_column_dend = FALSE        # Hides the global dendrogram (optional)
)
dev.off()

### violin plot
cell_types <- c("CT-L6","IT-L23","DG","ODC","CA1","CA3")
prop_df <- as.data.frame(t(cell_proportion))
prop_df$group <- Idents(P21_brain)  # add group label
prop_df <- prop_df %>% dplyr::select(c(cell_types,"group"))

long_df <- prop_df %>%
  tidyr::pivot_longer(
    cols = -group,
    names_to = "cell_type",
    values_to = "expression"
  )

ggplot(long_df, aes(x = factor(group), y = expression, fill = cell_type)) +
  geom_violin(scale = "width", trim = T, color = "black") +
  labs(
    x = "",
    y = "Proportion",
    fill = "Cell Type",
    title = ""
  ) + theme_bw()
```

```{r}
library(viridis)
spatial_plot <- function(partition, spatial_barcode) {
  set.seed(123)
  group <- partition
  spatial_coordinate <- read.table("~/SpM/spatial_barcodes.txt", sep = "\t")
  # Match spatial barcodes to coordinates
  idx <- match(spatial_barcode, spatial_coordinate$V4)
  spatial_coordinate_filter <- spatial_coordinate[idx, ]
  spatial_heatmap <- cbind(spatial_coordinate_filter$V2, spatial_coordinate_filter$V3, group)
  rownames(spatial_heatmap) <- spatial_barcode
  # Add empty (missing) barcodes
  empty_barcode_idx <- which(!spatial_coordinate$V4 %in% rownames(spatial_heatmap))
  empty_barcode <- cbind(spatial_coordinate[empty_barcode_idx, ]$V2,
                         spatial_coordinate[empty_barcode_idx, ]$V3,
                         rep(NA, length(empty_barcode_idx)))
  spatial_heatmap <- as.data.frame(rbind(spatial_heatmap, empty_barcode))
  colnames(spatial_heatmap) <- c("V1", "V2", "group")
  # Factorize and format
  spatial_heatmap$V1 <- as.factor(spatial_heatmap$V1)
  spatial_heatmap$V2 <- as.factor(spatial_heatmap$V2)
  # Check if group is numeric
  is_continuous <- is.numeric(group)
  # Convert group to appropriate type
  if (is_continuous) {
    spatial_heatmap$group <- as.numeric(spatial_heatmap$group)
  } else {
    spatial_heatmap$group <- factor(spatial_heatmap$group)
  }
  # Begin plotting
  p <- ggplot(spatial_heatmap, aes(x = V2, y = V1)) +
    geom_point(aes(color = group), size = 1.5, na.rm = TRUE) +
    theme_bw() +
    xlab("") + ylab("") +
    scale_x_discrete(limits = rev(levels(spatial_heatmap$V2))) +
    scale_y_discrete(limits = rev) +
    coord_fixed(ratio = 1) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank())
  # Apply appropriate color scale
  if (is_continuous) {
    p <- p + scale_color_viridis_c(na.value = "white", option = "viridis")
  } else {
    p <- p + scale_color_discrete(na.value = "white", na.translate = FALSE)
  }
  return(p)
}
```


### Human lymph node 

parallel -j 20 'biscuit vcf2bed -k 1 -c {} | awk -v OFS="\t" '\''{ print $1, $2+1, $3, $4, $5, $6 }'\'' > {.}.cov' ::: *.vcf.gz
scbs prepare *.cov ../../scbs/SpMETSLHLN4D

### use knee plot to determine minimum CG number for filter
### scbs filter --min-sites 63412 --min-meth 0 --max-meth 100 SpMETSLHLN4D SpMETSLHLN4D_filtered

scbs smooth SpMETSLHLN4D
scbs scan --threads 20 SpMETSLHLN4D SpMETSLHLN4D.bed
scbs matrix --threads 20 SpMETSLHLN4D.bed SpMETSLHLN4D SpMETSLHLN4D_matrix

```{r}
data_sample <- "SpMETRTHLN4RXT19"
path <- paste(paste("/mnt/isilon/zhou_lab/projects/20230531_Spatial_methseq/RNA/", data_sample, sep = '/'), "raw", sep = '/')
gene_matrix <- Read10X(data.dir = path)
assay = "Spatial"
filter.matrix = TRUE
slice = "slice1"
object <- CreateSeuratObject(counts = gene_matrix, assay = assay)
image <- Read10X_Image(image.dir = "/mnt/isilon/zhou_lab/projects/20230531_Spatial_methseq/Spatial/10_SpMETSLHLN4D_SpMETRTHLN4RXT19/spatial", filter.matrix = filter.matrix)
image <- image[Cells(x = object)]
DefaultAssay(object = image) <- assay
object[[slice]] <- image
problematic_cells <- which(object$nCount_Spatial == 0)
if (length(problematic_cells) > 0) {
  object <- subset(object, cells = -problematic_cells)
}
spatial.obj <- object
spatial.obj <- suppressWarnings(SCTransform(spatial.obj, assay = "Spatial",verbose=TRUE))
RNA_data <- spatial.obj@assays$SCT$scale.data
spatial.obj <- RunPCA(spatial.obj, assay = "SCT", verbose = FALSE)
spatial.obj <- FindNeighbors(spatial.obj, reduction = "pca", dims = 1:30,random.seed = 123)
spatial.obj <- FindClusters(spatial.obj, verbose = FALSE, resolution = 0.6,random.seed = 123)
spatial.obj <- RunUMAP(spatial.obj, reduction = "pca", dims = 1:30)

meth_mtx <- read.csv("~/tmp/Spatial_methseq/scbs/SpMETSLHLN4D_matrix/methylation_fractions.csv.gz", row.names=1) %>% as.matrix()
### tbl from yame summary
tbl$V8 <- as.numeric(tbl$V8)
barcodes_pass <- tbl %>% filter(V8 > 63412) %>% filter(V8 < 1100000) %>% select(V2)
barcodes_pass <- barcodes_pass[,1]
filter_barcode <- sub("\\..*$", "", rownames(meth_mtx))
rownames(meth_mtx) <- filter_barcode
DNAm_data <- meth_mtx[rownames(meth_mtx) %in% barcodes_pass,]
DNAm_data <- as.data.frame(DNAm_data)
spatial_barcode <- sub("\\..*$", "", rownames(DNAm_data))
spatial_coordinate <- read.table("~/SpM/spatial_barcodes.txt",sep="\t")
image_barcode <- rownames(spatial.obj@images$slice1@coordinates)
image_barcode <- str_replace_all(image_barcode, "C", "T")
image_barcode = paste0(substr(image_barcode,9,16), substr(image_barcode,1,8))
image_barcode <- image_barcode[which(image_barcode %in% spatial_barcode)]
idx <- match(image_barcode,spatial_coordinate$V4)
spatial_coordinate_filter <- spatial_coordinate[idx,]
idx <- match(spatial_coordinate_filter$V4,spatial_barcode)
DNAm_data <- DNAm_data[idx,]
spatial_barcode <- spatial_barcode[idx]
barcodes_to_keep <- sub("\\..*$", "", rownames(DNAm_data))
#saveRDS(barcodes_to_keep,"/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20251020_HLN4D_barcodesK.rds")
```

### imputation
```{r}
library(irlba)
# PCA that iteratively imputes missing values
imputeColMean <- function(mtx) {
  #na_percentage <- sapply(mtx, function(x) sum(is.na(x)) / length(x)) * 100
  #mtx <- mtx[, na_percentage < 20]
  k <- which(is.na(mtx), arr.ind=TRUE)
  mtx[k] <- colMeans(mtx, na.rm=TRUE)[k[,2]]
  mtx
}

prcomp_iterative <- function(x, n=10, n_iter=100, min_gain=0.01, ...) {
  mse <- rep(NA, n_iter)
  na_loc <- is.na(x)
  x <- imputeColMean(x)
  #x[na_loc] = 0  # 0 is our first guess
  for (i in 1:n_iter) {
    prev_imp <- x[na_loc]  # what we imputed in the previous round
    # PCA on the imputed matrix
    pr <- prcomp_irlba(x, center = T, scale. = F, n = n, ...)
    # impute missing values with PCA
    new_imp <- (pr$x %*% t(pr$rotation))[na_loc]
    x[na_loc] <- new_imp
    # compare our new imputed values to the ones from the previous round
    mse[i] = mean((prev_imp - new_imp) ^ 2)
    # if the values didn't change a lot, terminate the iteration
    gain <- mse[i] / max(mse, na.rm = T)
    if (gain < min_gain) {
      message(paste(c("\n\nTerminated after ", i, " iterations.")))
      break
    }
  }
  pr$mse_iter <- mse[1:i]
  list(pr,x)
}

pca <- DNAm_data %>%
  scale(center = T, scale = F) %>%
  prcomp_iterative(n = 30, n_iter = 100)
#saveRDS(pca,"~/tmp/20251020_HLD4N.rds")
```

```{r}
data_sample <- "SpMETRTHLN4RXT19"
path <- paste(paste("/mnt/isilon/zhou_lab/projects/20230531_Spatial_methseq/RNA/", data_sample, sep = '/'), "raw", sep = '/')
gene_matrix <- Read10X(data.dir = path)
bc_df <- data.frame(RNA_bc=colnames(gene_matrix), row.names = colnames(gene_matrix))
bc_df$DNA_bc <- str_replace_all(bc_df$RNA_bc, "C", "T")
rownames(bc_df) <- bc_df$DNA_bc

### from imputation
DNAm_data <- as.data.frame(pca[[2]])
DNAm_data <- DNAm_data[match(barcodes_to_keep,rownames(DNAm_data)),]
### scale this from 0 to 1
DNAm_data <- as.data.frame(lapply(DNAm_data, function(x) (x - min(x)) / (max(x) - min(x))))
DNAm_matrix <- DNAm_data
rownames(DNAm_matrix) <- barcodes_to_keep

rownames(DNAm_matrix) <- sub("\\..*$", "", rownames(DNAm_matrix))
rownames(DNAm_matrix) <- paste0(substr(rownames(DNAm_matrix),9,16), substr(rownames(DNAm_matrix),1,8))
bc_df <- bc_df[rownames(DNAm_matrix),]
all(rownames(bc_df)==rownames(DNAm_matrix))
rownames(DNAm_matrix) <- bc_df$RNA_bc
DNAm_matrix <- t(DNAm_matrix)

assay = "RNA"
filter.matrix = TRUE
slice = "slice1"
object <- CreateSeuratObject(counts = gene_matrix, assay = assay)

image <- Read10X_Image(image.dir = "/mnt/isilon/zhou_lab/projects/20230531_Spatial_methseq/Spatial/10_SpMETSLHLN4D_SpMETRTHLN4RXT19/spatial", filter.matrix = filter.matrix)
image <- image[Cells(x = object)]
DefaultAssay(object = image) <- assay
object[[slice]] <- image
spatial.obj <- object
### keep only barcodes that pass through two criterias
spatial.obj <- subset(spatial.obj, cells=colnames(DNAm_matrix))


spatial.obj[['DNAm']] <- CreateAssayObject(counts = DNAm_matrix)
DefaultAssay(spatial.obj) <- "DNAm"
VariableFeatures(spatial.obj) <- rownames(spatial.obj[["DNAm"]])
spatial.obj@assays$DNAm@scale.data <- as.matrix(spatial.obj@assays$DNAm@counts)

spatial.obj <- RunPCA(spatial.obj, assay = "DNAm", reduction.name = 'mpca', verbose = FALSE)

#spatial.obj <- readRDS("~/cell_type/20250430_human_spatial.rds")

spatial.obj <- FindNeighbors(spatial.obj, reduction = "mpca", dims = 1:30)
spatial.obj <- FindClusters(spatial.obj, verbose = FALSE, resolution = 1)
spatial.obj <- RunUMAP(spatial.obj, reduction = "mpca",  reduction.name = "meth.umap", dims = 1:30,n.neighbors = 20)


spatial.obj$DNAm_snn_res.1 <- paste0('D', spatial.obj$DNAm_snn_res.1)
n_clusters <- length(unique(spatial.obj$DNAm_snn_res.1))

cols <- c("#D51F26","#8A9FD1","#D8A767")

DimPlot(spatial.obj, reduction = 'meth.umap', repel = TRUE, group.by = 'DNAm_snn_res.1',pt.size = 0.7,alpha = 0.7,cols=cols) + NoGrid() + ggtitle("")
ggsave("~/figures/fuh1/20250724_HLN4_VMR_umap.pdf",width=3,height=3)

spatial.obj_barcode <- sub("\\..*$", "", colnames(spatial.obj))
spatial.obj_barcode <- paste0(substr(spatial.obj_barcode,9,16), substr(spatial.obj_barcode,1,8))

digits <- as.integer(gsub("D", "", spatial.obj@meta.data$DNAm_snn_res.1))
valid_barcodes <- readRDS("/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20251020_HLN4D_barcodesK.rds")

spatial_plot(as.factor(digits),valid_barcodes) + scale_color_manual(values=cols,na.value="white",na.translate = FALSE,labels = c("D0", "D1", "D2","D3")) + labs(color ="")
ggsave("~/figures/fuh1/20250724_HLN4_VMR.pdf",width=4.5,height=4)
```


### VMR and MethScope clustering

```{r}
library(Seurat)
library(ArchR)
reference_pattern <- system.file("extdata", "Zhou2025_HumanAtlas.cm", package = "MethScope")
#input_pattern <- GenerateInput("/home/fuh1/zhou_lab/projects/20230531_Spatial_methseq/pileup/SpMETSLHLN4D/SpMETSLHLN4D.cg",reference_pattern)
valid_barcodes <- readRDS("/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20251020_HLN4D_barcodesK.rds")

input_pattern <- readRDS("~/cell_type/20250423_HLN4_input.rds")
input_pattern <- input_pattern[rownames(input_pattern) %in% valid_barcodes,]
input_pattern <- input_pattern %>% select(-Pna)
#saveRDS(input_pattern,"~/cell_type/20250423_HLN4_input.rds")

reference_input <- GenerateInput("/mnt/isilon/zhou_lab/projects/20230727_all_public_WGBS/hg38/2025_ZhouPseudo.cg",reference_pattern)

input_pattern_imputed <- imputeRowMean(input_pattern,na_percent = 30)
reference_input_imputed <- imputeRowMean(reference_input,na_percent = 30)

# unique_patterns <- sortedFrequency_df %>% filter(unique == "Unique hypomethylated") %>% select(pattern)
# input_pattern_imputed_filter <- input_pattern_imputed[,unique_patterns[,1]]
# reference_input_imputed_filter <- reference_input_imputed[,unique_patterns[,1]]
# cell_proportion_filter <- nnls_deconv(reference_input_imputed_filter,input_pattern_imputed_filter,number_patterns = 62)

cell_proportion <- nnls_deconv(reference_input_imputed,input_pattern_imputed)
input_pattern_imputed <- cbind(input_pattern_imputed,t(cell_proportion))

# all_na_rows <- apply(input_pattern, 1, function(x) all(is.na(x)))
# input_pattern_filtered <- input_pattern[!all_na_rows, ]
# input_pattern_imputed <- imputeRowMean(input_pattern_filtered,na_percent = 40)

input_pattern_imputed <- imputeRowMean(input_pattern,na_percent = 1)
#input_pattern_imputed <- input_pattern_imputed %>% select(-P1,-P2)
lymph_spatial <- CreateSeuratObject(counts = t(input_pattern_imputed), assay = "DNAm")
VariableFeatures(lymph_spatial) <- rownames(lymph_spatial[['DNAm']])
DefaultAssay(lymph_spatial) <- "DNAm"
lymph_spatial <- NormalizeData(lymph_spatial, assay = "DNAm", verbose = FALSE)
lymph_spatial <- ScaleData(lymph_spatial, assay = "DNAm", verbose = FALSE)
#lymph_spatial@assays$DNAm@layers$scale.data <- as.matrix(lymph_spatial@assays$DNAm@layers$counts)

lymph_spatial <- RunPCA(lymph_spatial,assay="DNAm",reduction.name = 'mpca', verbose = FALSE)
lymph_spatial <- FindNeighbors(lymph_spatial, reduction = "mpca", dims = 1:30)
lymph_spatial <- FindClusters(lymph_spatial, verbose = FALSE, resolution = 0.7)
lymph_spatial <- RunUMAP(lymph_spatial, reduction = "mpca",  reduction.name = "meth.umap", dims = 1:30,n.neighbors = 20, min.dist = 0.3,spread = 0.3,set.op.mix.ratio = 0.3)

custom_colors <- c("#8A9FD1","#FEE500","#89C75F","#D51F26")
lymph_spatial$DNAm_snn_res.0.7 <- paste0("D",lymph_spatial$DNAm_snn_res.0.7)
DimPlot(lymph_spatial, reduction = 'meth.umap', repel = TRUE, group.by = 'DNAm_snn_res.0.7', pt.size = 0.7,alpha = 0.7,cols=custom_colors)  + NoGrid() + ggtitle("")
ggsave("~/figures/fuh1/20250724_HLN4_MethScope_umap.pdf",width=3,height=3)

partition <- Idents(lymph_spatial)
#prediction_result <- PredictCellType(Zhou2025_HumanAtlas_P1000,input_pattern,smooth = T,KNeighbor = 2)
#partition <- P21_brain$MethScope

spatial_barcode <- colnames(lymph_spatial)
spatial_plot(partition,spatial_barcode)  + scale_color_manual(values=custom_colors,na.value="white",na.translate = FALSE,labels = c("D0","D1","D2","D3","D4")) + labs(color ="")
ggsave("~/figures/fuh1/20250724_HLN4_MethScope.pdf",width=4.5,height=4)


cell_type <- "Tnaive_CD8"
test <- cell_proportion[cell_type,]
#test[which(test > 0.2)] <- 0.2
spatial_plot(test,names(cell_proportion[cell_type,]))+labs(color = cell_type)
#spatial_plot(cell_proportion_z_clean[cell_type,],names(cell_proportion_z_clean[cell_type,]))+labs(color = cell_type)


ggsave(paste0("~/figures/fuh1/20250501_HLN4_",cell_type,".pdf"),width=4.5,height=4)

```

```{R}
library(pheatmap)
cell_proportion_z <- t(scale(t(cell_proportion), center = TRUE, scale = TRUE))
cell_proportion_z_clean <- cell_proportion_z[!apply(cell_proportion_z, 1, function(x) all(is.na(x))), ]
cell_proportion_z_clean <- cell_proportion_z_clean[apply(cell_proportion_z_clean, 1, function(x) all(abs(x) <= 20)), ]
cell_proportion_z_clean <- pmax(pmin(cell_proportion_z_clean, 3), -3)

cell_type <- c("B_Mem","B_Naive","B_Plasma","Macrophage","NK_CD56","Tmem_CD4","Tmem_CD8","Tnaive_CD4","Tnaive_CD8")
cell_proportion_z_clean <- cell_proportion_z_clean[cell_type,]

prop_mat <- cell_proportion_z_clean
# Reorder samples by cluster
cluster_vec <- Idents(lymph_spatial)
cluster_order <- order(cluster_vec)
prop_ordered <- prop_mat[, cluster_order]  # reorder columns
clusters <- cluster_vec[cluster_order]

# Optional: assign a color per cluster
cluster_levels <- unique(clusters)
cluster_colors <- setNames(c("#8A9FD1","#FEE500","#89C75F","#D51F26","#D8A767"), cluster_levels)

# Prepare annotation for columns (samples)
annotation_col <- data.frame(Cluster = clusters)
rownames(annotation_col) <- colnames(prop_ordered)

# Plot the heatmap
pdf("~/figures/fuh1/20250502_lymph_proportion.pdf",width=6,height=4.5)
pheatmap(prop_ordered,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         annotation_col = annotation_col,
         annotation_colors = list(Cluster = cluster_colors),
         show_colnames = FALSE,
         fontsize_row = 10,
         scale = "none",
         main = "",
         color = hcl.colors(50, "BluYl"))
dev.off()

### cluster columns 

# Prepare annotation for columns
cluster_levels <- unique(clusters)
cluster_colors <- setNames(
  c("#8A9FD1","#FEE500","#89C75F","#D51F26","#D8A767"),
  cluster_levels
)

# Annotation dataframe for columns
annotation_col <- data.frame(Cluster = clusters)
rownames(annotation_col) <- colnames(prop_ordered)

# Define annotation with colors
col_ha <- HeatmapAnnotation(
  df = annotation_col,
  col = list(Cluster = cluster_colors),
  show_annotation_name = TRUE
)

library(ComplexHeatmap)
min_val <- min(prop_ordered)
max_val <- max(prop_ordered)
# Get all cluster levels you want, in order
desired_order <- 0:4
clusters_named <- factor(
  clusters,
  levels = desired_order,
  labels = paste0("Cluster ", desired_order)
)

pdf("~/figures/fuh1/20250707_lymph_proportion.pdf",width=6,height=4.5)
Heatmap(
  as.matrix(prop_ordered),
  name = "Proportion",
  top_annotation = col_ha,
  cluster_rows = FALSE,
  cluster_columns = TRUE,         # This will cluster within each cluster
  column_split = clusters_named,        # Split by cluster group
  column_title = NULL,   
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 10),
  col = colorRamp2(
  seq(min_val, max_val, length.out = 50),
  hcl.colors(50, "Inferno")
  ),
  show_column_dend = FALSE        # Hides the global dendrogram (optional)
)
dev.off()
```

### UXM comparison

./uxm deconv ~/tmp/wgbs_tools/SpMETSLHLN4D/*pat.gz --atlas ./supplemental/Atlas.U25.l4.hg38.tsv -o 20250820_SpMETSLHLN4D_UXM.csv

mkdir -p results

find ~/tmp/wgbs_tools/SpMETSLHLN4D -name '*.pat.gz' | parallel -j 8 '
  f={};
  base=$(basename "$f" .pat.gz);
  ./uxm deconv "$f" \
    --atlas ~/tmp/UXM_deconv/supplemental/Atlas.U25.l4.hg38.full.tsv \
    -o results_full/${base}.csv \
    2> results_full/${base}.log || true
'

```{r}
library(tidyverse)

files <- list.files("~/tmp/UXM_deconv/results_full", pattern = "*.csv", full.names = TRUE)
merged <- files %>%
  map(function(f) {
    sample <- tools::file_path_sans_ext(basename(f))
    read_csv(f, show_col_types = FALSE) %>%
      rename(!!sample := 2)
  }) %>%
  reduce(full_join, by = "CellType")
zero_cols <- sapply(merged[,-1], function(x) all(x == 0))
merged_nonzero <- merged[,-1]
merged_nonzero <- merged_nonzero[,zero_cols==F]
rownames(merged_nonzero) <- merged$CellType

valid_barcodes <- readRDS("/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20251020_HLN4D_barcodesK.rds")


idx <- match(valid_barcodes,colnames(merged[,-1]))
df <- merged[,-1]
idx <- idx[!is.na(idx)]  
df <- df[,idx]

rownames(df) <- merged$CellType

cell_type <- "Blood-B"
test <- as.numeric(as.data.frame(df[cell_type,])[1,])
spatial_plot(test,colnames(df))+labs(color = cell_type)
ggsave(paste0("~/figures/fuh1/20250820_HLN4_",cell_type,"_uxm.pdf"),width=4.5,height=4)
# test <- read.table("~/tmp/UXM_deconv/supplemental/Atlas.U25.l4.hg38.full.tsv",header = T)
# test2 <- read.table("~/tmp/UXM_deconv/supplemental/Atlas.U25.l4.hg38.tsv",header=T)
```


### Human Skin

```{r}
library(Seurat)
library(ArchR)
reference_pattern <- system.file("extdata", "Zhou2025_HumanAtlas.cm", package = "MethScope")
input_pattern <- GenerateInput("/home/fuh1/zhou_lab/projects/20230531_Spatial_methseq/pileup/SpMETSLHSK08DD/SpMETSLHSK08DD.cg",reference_pattern)
#valid_barcodes <- readRDS("/home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20250430_HLN4D_barcodesK.rds")


input_pattern <- input_pattern[rownames(input_pattern) %in% image_barcode,]
input_pattern <- input_pattern %>% select(-Pna)
#saveRDS(input_pattern,"~/cell_type/20250423_HLN4_input.rds")

prediction_result <- PredictCellType(training_model,input_pattern,smooth = F)  


reference_input <- GenerateInput("/mnt/isilon/zhou_lab/projects/20230727_all_public_WGBS/hg38/2025_ZhouPseudo.cg",reference_pattern)

input_pattern_imputed <- imputeRowMean(input_pattern,na_percent = 30)
reference_input_imputed <- imputeRowMean(reference_input,na_percent = 30)

# unique_patterns <- sortedFrequency_df %>% filter(unique == "Unique hypomethylated") %>% select(pattern)
# input_pattern_imputed_filter <- input_pattern_imputed[,unique_patterns[,1]]
# reference_input_imputed_filter <- reference_input_imputed[,unique_patterns[,1]]
# cell_proportion_filter <- nnls_deconv(reference_input_imputed_filter,input_pattern_imputed_filter,number_patterns = 62)

cell_proportion <- nnls_deconv(reference_input_imputed,input_pattern_imputed)
input_pattern_imputed <- cbind(input_pattern_imputed,t(cell_proportion))

# all_na_rows <- apply(input_pattern, 1, function(x) all(is.na(x)))
# input_pattern_filtered <- input_pattern[!all_na_rows, ]
# input_pattern_imputed <- imputeRowMean(input_pattern_filtered,na_percent = 40)

input_pattern_imputed <- imputeRowMean(input_pattern,na_percent = 30)
input_pattern_imputed <- input_pattern_imputed %>% select(-P1,-P2)

### image barcode
# input_pattern_imputed <- input_pattern_imputed[rownames(input_pattern_imputed) %in% image_barcode,]
skin_spatial <- CreateSeuratObject(counts = t(input_pattern_imputed), assay = "DNAm")

assay = "DNAm"
filter.matrix = TRUE
slice = "slice1"
image <- Read10X_Image(image.dir = "/mnt/isilon/zhou_lab/projects/20230531_Spatial_methseq/Spatial/SpMETSLHSK08D/", filter.matrix = filter.matrix)

image_barcodes <- rownames(image@coordinates)
image_barcodes_fixed <- image_barcodes |>
  str_replace_all("C", "T") |>
  (\(x) paste0(substr(x, 9, 16), substr(x, 1, 8)))()
rownames(image@coordinates) <- image_barcodes_fixed

image <- image[Cells(x = skin_spatial)]
DefaultAssay(skin_spatial = image) <- assay
skin_spatial[[slice]] <- image

image_barcode <- rownames(skin_spatial@images$slice1@coordinates)

VariableFeatures(skin_spatial) <- rownames(skin_spatial[['DNAm']])
DefaultAssay(skin_spatial) <- "DNAm"
skin_spatial <- NormalizeData(skin_spatial, assay = "DNAm", verbose = FALSE)
skin_spatial <- ScaleData(skin_spatial, assay = "DNAm", verbose = FALSE)
#lymph_spatial@assays$DNAm@layers$scale.data <- as.matrix(lymph_spatial@assays$DNAm@layers$counts)

skin_spatial <- RunPCA(skin_spatial,assay="DNAm",reduction.name = 'mpca', verbose = FALSE)
skin_spatial <- FindNeighbors(skin_spatial, reduction = "mpca", dims = 1:30)
skin_spatial <- FindClusters(skin_spatial, verbose = FALSE, resolution = 0.7)
skin_spatial <- RunUMAP(skin_spatial, reduction = "mpca",  reduction.name = "meth.umap", dims = 1:30,n.neighbors = 20)

custom_colors <- c("#8A9FD1","#FEE500","#89C75F","#D51F26","#FFA500","#00FFF0")
skin_spatial$DNAm_snn_res.0.7 <- paste0("D",skin_spatial$DNAm_snn_res.0.7)
DimPlot(skin_spatial, reduction = 'meth.umap', repel = TRUE, group.by = 'DNAm_snn_res.0.7', pt.size = 0.7,alpha = 0.7,cols=custom_colors)  + NoGrid() + ggtitle("")
ggsave("~/figures/fuh1/20251002_HSK08D_MethScope_umap.pdf",width=3,height=3)

partition <- Idents(skin_spatial)
prediction_result <- PredictCellType(Zhou2025_HumanAtlas_P1000,input_pattern,smooth = F)
prediction_result <- prediction_result[match(colnames(skin_spatial),rownames(prediction_result)),]
partition <- prediction_result$prediction_label
partition <- ifelse(partition == "NK CD56",1,0)

spatial_barcode <- colnames(skin_spatial)
spatial_plot(partition,spatial_barcode) + theme(legend.position = "None") + ggtitle("NK CD56")


  scale_color_discrete(
    labels = c("Alpha cells", "ASC cells", "Type I AT", "Type II AT", "Memory B", "Naive B")
  ) #+ scale_color_manual(values=custom_colors,na.value="white",na.translate = FALSE,labels = c("D0","D1","D2","D3","D4")) + labs(color ="")
ggsave("~/figures/fuh1/20251003_HSK08D_MethScope.pdf",width=4.5,height=4)



cell_type <- "SST"
test <- cell_proportion[cell_type,]
spatial_plot(test,names(cell_proportion[cell_type,]))+labs(color = cell_type)


