---
title: "20250130_patterns"
output: html_document
date: "2025-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Pattern frequency and enrichment

```{r}
#2025_ZhouPseudo_binstring
reference_set = readLines("~/tmp/20240322_SparseM/2021_LiuMajorTypePseudo_binstring")
#Calculate the frequency of each string
stringFrequency <- table(reference_set)
frequencyDF <- as.data.frame(stringFrequency)
#Sort the data frame by frequency
sortedFrequency <- frequencyDF[order(-frequencyDF$Freq),]
sortedFrequency$pattern <- paste0("P",seq(1:nrow(sortedFrequency)))
count1s <- sapply(as.character(sortedFrequency$reference_set),function(x)sum(strsplit(x, NULL)[[1]] == "0"))
one_one <- sortedFrequency[which(count1s==1),]
idx_file <- read.table("/mnt/isilon/zhou_lab/projects/20230727_all_public_WGBS/mm10/2021_LiuMajorTypePseudo.cg.idx",sep="\t")
row_name <- sapply(as.character(one_one$reference_set),function(x)idx_file$V1[which(strsplit(x, NULL)[[1]] == "0")])
one_one$cell_type <- row_name
  
data_visualize <- one_one
data_visualize$reference_set <- factor(data_visualize$reference_set,levels = data_visualize$reference_set)
data_visualize$cell_type <- factor(data_visualize$cell_type,levels = data_visualize$cell_type)

ggplot(data_visualize, aes(x=cell_type, y=Freq)) + 
  geom_bar(stat = "identity")+ 
  theme_bw() + theme(axis.title.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 70, vjust = 1, hjust=1))+ggtitle("Liu 2021 Major Cell Type Hypomethylation Patterns")+labs(y="#CGs")
ggsave("~/figures/fuh1/20250212_Liu2021_major_hypo.pdf",width=6,height=4)


```

### Visualize cumulative CpG along the patterns

```{r}
reference_set = readLines("~/tmp/20240322_SparseM/2021_LiuMajorTypePseudo_binstring")
stringFrequency <- table(reference_set)
frequencyDF <- as.data.frame(stringFrequency)
sortedFrequency <- frequencyDF[order(-frequencyDF$Freq),]
sortedFrequency$pattern <- paste0("P",seq(1:nrow(sortedFrequency)))

sortedFrequency$Cumulative_Freq = cumsum(sortedFrequency$Freq)
sortedFrequency_df <- sortedFrequency[1:1000,]

count1s <- sapply(as.character(sortedFrequency_df$reference_set),function(x)sum(strsplit(x, NULL)[[1]] == "1"))
count0s <- sapply(as.character(sortedFrequency_df$reference_set),function(x)sum(strsplit(x, NULL)[[1]] == "0"))

sortedFrequency_df$unique <- "Not unique"
sortedFrequency_df$unique[which(count1s==1)] <- "Unique hypermethylated"
sortedFrequency_df$unique[which(count0s==1)] <- "Unique hypomethylated"

sortedFrequency_df$size <- ifelse(sortedFrequency_df$unique == "Not unique",0,1)
sortedFrequency_df$pattern_number <- 1:1000

# cumulative_list <- list()
# unique_classes <- unique(sortedFrequency_df$unique)
# for (class in unique_classes) {
#     # Filter rows for the current class
#     temp_df <- sortedFrequency_df %>%
#         filter(unique == class) %>%
#         arrange(pattern_number) %>%  # Sort by pattern
#         mutate(Cumulative_Freq = cumsum(Freq))  # Compute cumulative frequency
#     # Store the result in the list
#     cumulative_list[[class]] <- temp_df
# }
# sortedFrequency_df_unique <- do.call(rbind, cumulative_list)


ggplot(sortedFrequency_df, aes(x = pattern_number, y = Cumulative_Freq, group = 1, 
               color = unique,size=size)) + 
  geom_line(color = "#F1AEA7", linewidth = 1) +  # Line for cumulative trend
  geom_point(alpha = 1) + # Points with color and size mapping
  scale_y_log10() +                       # Log10 scale for y-axis
  scale_color_manual(values = c("#A6DAEF", "#FDD379", "#B0D9A5")) +  # Set distinct colors
  scale_size_continuous(range = c(1.5, 2)) + # Adjust point size scale
  labs(title = "", 
       x = "Pattern", 
       y = "Cumulative CpG #", 
       color = "Status", 
       size = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text = element_text(size = 11)) + guides(size = "none")


ggsave("~/figures/fuh1/20240206_cumulative_CGs_Liu.pdf",width=5,height=4)

```

### Feature importance
```{r}
Liu2021_MouseBrain_model <- Liu2021_MouseBrain_P1000
importance_matrix <- xgb.importance(model = Liu2021_MouseBrain_model)
num_classes <- Liu2021_MouseBrain_model$params$num_class
num_trees <- Liu2021_MouseBrain_model$niter

# Compute how many trees per class
trees_per_class <- num_trees / num_classes
# Extract feature importance for each class
feature_importance_by_class <- list()
for (class_index in 1:num_classes) {
  # Identify trees belonging to the class
  class_trees <- seq(class_index, num_trees, by = num_classes)-1
  # Extract importance for these trees
  feature_importance_by_class[[paste0("Class_", class_index)]] <- xgb.importance(model = Liu2021_MouseBrain_model, trees = class_trees)
}

# Convert to data frame for plotting
class_13_importance <- feature_importance_by_class[["Class_13"]]

# singe class importance Plot
ggplot(class_13_importance, aes(x = reorder(Feature, Gain), y = Gain)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Feature Importance for DG", x = "Feature", y = "Gain")

top_features <- data.frame(Class = character(), Feature = character(), Gain = numeric(), stringsAsFactors = FALSE)
for (class_name in names(feature_importance_by_class)) {
  class_importance <- feature_importance_by_class[[class_name]]
  top_feature <- class_importance[which.max(class_importance$Gain), ]
  top_features <- rbind(top_features, data.frame(Class = class_name, Feature = top_feature$Feature, Gain = top_feature$Gain))
}
top_features$Class <- Liu2021_MouseBrain_model$cell_type
top_features$Class <- factor(top_features$Class, levels = top_features$Class)

### run the above for classifying which feature is unique 

top_features$Unique <- sortedFrequency_df$unique[match(top_features$Feature,sortedFrequency_df$pattern)]

ggplot(top_features, aes(x = Class, y = Gain, fill = Unique)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Feature),angle = 90,vjust = 0.5,hjust=1, size = 3, color = "white") +  # Annotate feature names
  labs(title = "",
       x = "",
       y = "Gain (Feature Importance)",
       fill = "") +
  theme_bw() +
  theme(legend.position = "top",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 70, vjust = 1, hjust = 1)) + scale_fill_brewer(palette = "Set2")
ggsave("~/figures/fuh1/20250212_Liu_featureimportance.pdf",width=6,height=4)
```

### KYCG enrichment of each pattern

```{r}
pattern <- which(sortedFrequency$Freq == 50057)
sortedFrequency[which(sortedFrequency$Freq <= 50),]$pattern <- "Pna"
patterns_out <- sortedFrequency$pattern[match(reference_set, sortedFrequency$reference_set)]
patterns_out_enrich <- ifelse(patterns_out == "P23",1,0)
write.table(patterns_out_enrich,"~/zhoulab/labprojects/20231112_Hongxiang/20250130_LiuCA3.txt",quote=FALSE,col.names = FALSE,row.names = FALSE)
```

yame pack ~/zhoulab/labprojects/20231112_Hongxiang/20250130_LiuCA3.txt -fb ~/zhoulab/labprojects/20231112_Hongxiang/20250130_LiuCA3.cg

while IFS= read -r cm_file; do yame summary -m "$cm_file" /home/fuh1/zhoulab/labprojects/20231112_Hongxiang/20250130_LiuCA3.cg >> /home/fuh1/intermediate/kycg/20250130_LiuCA3_hypo; done < ~/intermediate/kycg/20241219_KYCG_mm10.txt

```{r}
enrichment <- read.table("/home/fuh1/intermediate/kycg/20250130_LiuCA3_hypo",header=T)
enrichment <- enrichment %>% filter(QFile != "QFile")
enrichment <- enrichment %>% filter(is.na(Mask) == F)
enrichment$Log2OddsRatio <- as.numeric(enrichment$Log2OddsRatio)
enrichment$N_overlap <- as.numeric(enrichment$N_overlap)
enrichment <- enrichment %>% filter(Log2OddsRatio != Inf)
enrichment <- enrichment %>% filter(Log2OddsRatio != -Inf)
enrichment <- enrichment %>% filter(N_overlap >= 20)
enrichment <- enrichment[order(enrichment$Log2OddsRatio,decreasing = T),]
p_value_data <- as.data.frame(enrichment)[,c(7,6,8,5)]
p_values <- apply(p_value_data,1,function(x){testEnrichmentFisherN(as.numeric(x[1]),as.numeric(x[2]),as.numeric(x[3]),as.numeric(x[4]))$p.value})

enrichment$P.value <- p.adjust(p_values,method="fdr")

plot_results <- enrichment[,c("MFile","Mask","Log2OddsRatio","P.value")]

plot_results$dot_size <- -log10(plot_results$P.value) 

top_terms <- plot_results %>%
  group_by(MFile) %>%
  top_n(n = 5, wt = Log2OddsRatio) %>%
  ungroup()

extract_terms <- function(x) {
  parts <- strsplit(x, ";")[[1]]
  if (length(parts) > 2 && tail(parts, n = 1) == "_") {
    # If the last term is "_", take the second-to-last term
    return(paste(parts[1], parts[length(parts) - 1], sep = ";"))
  } else if (length(parts) > 1) {
    # Otherwise, take the last term
    return(paste(parts[1], tail(parts, n = 1), sep = ";"))
  } else {
    return(x) # Return as-is if the string doesn't have a delimiter
  }
}
top_terms$Mask <- sapply(top_terms$Mask, extract_terms)
top_terms <- top_terms %>%
  group_by(MFile) %>%
  arrange(desc(dot_size)) %>%
  slice_head(n = 5) %>%
  ungroup()
top_terms <- top_terms[order(top_terms$Log2OddsRatio),]

top_terms$Mask <- factor(top_terms$Mask,levels=unique(top_terms$Mask))

top_terms$dot_size[which(top_terms$dot_size > 100)] = 100

top_terms$MFile <- gsub(".*","",top_terms$MFile)

ggplot(top_terms, aes(x = Log2OddsRatio, y = Mask, color = MFile, size = dot_size)) +
  geom_point(alpha = 0.7) + # Alpha adjusts transparency
  theme_minimal() +
  labs(
    x = "Log2 Odds Ratio",
    y = "",
    title = "",
    color = "",
    size = "-log10(FDR)"
  ) +
  scale_size_continuous(range = c(2, 10)) + # Adjust dot size range
  theme(
    axis.text.y = element_text(size = 10, hjust = 1),
    axis.title = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size = 14)
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  xlim(-3,3) + scale_color_brewer(palette = "Set2")# Add vertical line at 0 
ggsave("~/figures/fuh1/20250130_CA3Liu_hypo.pdf",width=6,height=5)

```

### KYCG enrichment pattern of TFBS

```{r}
pattern_list <- top_features %>% filter(Unique == "Unique hypomethylated")
pattern_list_enrich <- pattern_list$Feature

sortedFrequency[which(sortedFrequency$Freq <= 50),]$pattern <- "Pna"
patterns_out <- sortedFrequency$pattern[match(reference_set, sortedFrequency$reference_set)]
patterns_out_enrich_full <- c()
for (i in 1:length(pattern_list_enrich)){
  patterns_out_enrich <- ifelse(patterns_out == pattern_list_enrich[i],1,0)
  patterns_out_enrich_full <- cbind(patterns_out_enrich_full,patterns_out_enrich)
}

for (i in 1:ncol(patterns_out_enrich_full)){
  write.table(patterns_out_enrich_full[,i],"~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern.txt",quote=FALSE,col.names = FALSE,row.names = FALSE)
  text <- paste0("yame pack ~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern.txt -fb ~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern/",i,".cg")
  system(text)
}
# mergeCG2 ~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern/ ~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern.cg
# yame summary -m /home/fuh1/references/mm10/KYCGKB_mm10/TFBSrm.20221005.cm ~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern.cg > ~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern_TFBS

library(dplyr)
library(tidyr)
TFBS_result <- read.table("~/zhoulab/labprojects/20231112_Hongxiang/20250409_Liu_pattern_TFBS",header=T)
TFBS_result$Log2OddsRatio <- as.numeric(TFBS_result$Log2OddsRatio)
# Step 1: Identify top 5 masks globally based on average Log2OddsRatio
top_masks <- TFBS_result %>%
  group_by(Query) %>%
  arrange(Query, dplyr::desc(Log2OddsRatio)) %>%   # Use explicit namespace
  slice_head(n = 2) %>%
  select(Query,Mask)

# Step 2: Filter original data to include only those top 5 masks
filtered_df <- TFBS_result %>%
  filter(Mask %in% top_masks$Mask)

# Step 3: Transform to wide matrix (Query × Mask)
matrix_df <- filtered_df %>%
  select(Query, Mask, Log2OddsRatio) %>%
  pivot_wider(names_from = Mask, values_from = Log2OddsRatio)

matrix_df <- as.data.frame(matrix_df)
rownames(matrix_df) <- paste0(pattern_list_enrich[matrix_df$Query]," (", pattern_list$Class[matrix_df$Query], ")")
matrix_df <- matrix_df[,-1]
df_matrix_clean <- matrix_df
df_matrix_clean[df_matrix_clean == -Inf] <- NA
df_zscore <- as.data.frame(scale(df_matrix_clean, center = TRUE, scale = TRUE))


df_zscore$Query <- rownames(df_zscore)
df_long <- df_zscore %>%
  pivot_longer(
    -Query,
    names_to = "Mask",
    values_to = "Zscore"
  )

ggplot(df_long, aes(x = Mask, y = Query, fill = Zscore)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "viridis", na.value = "white") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Transcription Factor Binding Sites",
    y = "",
    fill = "Z-score"
  )
ggsave("~/figures/fuh1/20250409_TFBS_liu.pdf",width=7,height=5)
```


### Sparsity levels across different datasets
```{r}
suppressMessages(library(readr))
df1 = read_tsv("~/zhou_lab/projects/20230531_Spatial_methseq/features/SpMETSLE14DM/ChromHMM.20220414_new") %>% dplyr::filter(!is.na(Mask)) 
df2 = read_tsv("~/zhou_lab/projects/20230531_Spatial_methseq/features/SpMETSLB14DM/ChromHMM.20220414_new") %>% dplyr::filter(!is.na(Mask))
df0 = read_tsv("~/zhoulab/labprojects/20210610_zhouw3/20230804_public_mm10_chromHMM_merged.tsv.gz") %>% dplyr::filter(!is.na(Mask))
df0$File = gsub("_mm10","",df0$File)
df1$File = "Spatial E11 Embryo (50 \u03BCm)"
df2$File = "Spatial P21 Brain (20 \u03BCm)"
df0 = bind_rows(df0,df1,df2)
df0 = df0 %>% dplyr::select(File, QFile, N_query) %>% unique

grDevices::cairo_pdf("~/figures/fuh1/SparseInfer/202401007_NCG_covered.pdf",width=5, height=5)
ggplot(df0) + geom_hline(yintercept=21000000) + geom_hline(yintercept=21000000*0.1, lty="dashed") + geom_hline(yintercept=21000000*0.01, lty="dotted") + scale_y_continuous(trans="log10") + geom_violin(aes(File, N_query, fill = File),trim=TRUE) + ylab("#CpGs covered") + xlab("") + theme_bw()+ theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  panel.background = element_blank()
) + theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust=1))+ theme(text = element_text(size = 11), legend.position = "None") + coord_flip() 
dev.off()
```

### Number of patterns in 5 cross-validation
```{r}
suppressMessages(library(stringr))
suppressMessages(library(doParallel))
suppressMessages(library(foreach))
suppressMessages(library(xgboost))
suppressMessages(library(dplyr))
suppressMessages(library(caret))
Liu_rf_input <- readRDS("~/tmp/20240322_SparseM/20240322_Liu_training.RDS")
pattern <- lapply(1:12, function(i) 1:(2^i))
train <- Liu_rf_input[,-c(1,ncol(Liu_rf_input))]
cell_type <- Liu_rf_input$cell_type
column_order <- order(as.numeric(str_extract(colnames(train), "\\d+")))
train <- train[, column_order]
cell_type <- as.factor(cell_type)
cell_type <- as.numeric(as.factor(cell_type)) - 1
train = do.call(cbind, lapply(train[,1:5000], as.numeric))

numberOfClasses <- length(unique(cell_type))
xgb_params <- list("objective" = "multi:softprob",
                   "eval_metric" = "mlogloss",
                   "num_class" = numberOfClasses)
nround    <- round(sqrt(nrow(train)))
cv.nfold  <- 5
registerDoParallel(cores=12)

accuracy_rate <- foreach(i=1:12, .combine='c') %do% {
  train_selected <- train[,unlist(pattern[i])]
  dtrain <- xgb.DMatrix(data = train_selected, label= cell_type)
  cv_model <- xgb.cv(params = xgb_params,
                   data = dtrain, 
                   nrounds = nround,
                   nfold = cv.nfold,
                   verbose = FALSE,
                   prediction = TRUE)
  OOF_prediction <- data.frame(cv_model$pred) %>%
  mutate(max_prob = max.col(., ties.method = "last"),
         label = cell_type + 1)
  confusion_matrix <- confusionMatrix(factor(OOF_prediction$max_prob),
                factor(OOF_prediction$label),
                mode = "everything")
  confusion_matrix <- confusion_matrix$table
  accuracy_rate <- 1-(sum(confusion_matrix) - sum(diag(confusion_matrix))) / sum(confusion_matrix)
  accuracy_rate
}

pdf("~/figures/fuh1/SparseInfer/20241009_xgb_numberPatterns.pdf", width=5, height=5, onefile=FALSE)
plot(accuracy_rate*100,type = "b", pch = 18,xlab=expression(paste("2"^"x")~"number of patterns"),ylab="CV accuracy %",cex.lab=1.3, cex.axis=1.3, ylim = c(0,100))
abline(h = 95, col="red", lwd=1, lty=2)
dev.off()
```

### Number of patterns and sparsity levels

rm -f ~/tmp/20231122_downsample/bycell/*
doawk() { awk -v number=$1 -v seed=$2 -v gsm=$3 'NR>1 && int(substr($4,2)) <= 2000{print number, seed, gsm, $4,$10;}' $4;}
export -f doawk
cat ~/cell_type/20240322_Liu2021_testing_sample.txt | awk '{for(i=10;i<=19;++i) {print 2**i,$1;}}' | parallel --col-sep '\t' -j 48 ':>~/tmp/20231122_downsample/bycell/{1}_{2}.gz; for seed in {1..1}; do yame subset /mnt/isilon/zhou_lab/projects/20220324_SingleCellMeth/mm10/2021_Liu.cg {2} | yame dsample -s $seed -N {1} - | yame summary -m ~/cell_type/Tool/MethScope/inst/extdata/Liu2021_MouseBrain.cm - | doawk {1} $seed {2} - | gzip -c >>~/tmp/20231122_downsample/bycell/{1}_{2}.gz; done'
find ~/tmp/20231122_downsample/bycell/ -name '*.gz' | xargs cat >~/tmp/20231122_downsample/dsample.tsv.gz

yame subset /mnt/isilon/zhou_lab/projects/20220324_SingleCellMeth/mm10/2021_Liu.cg -l ~/cell_type/20240322_Liu2021_training_sample.txt -o ~/intermediate/20240322_Liu2021_training.cg

```{r}
library(readr)
library(tidyr)
library(readxl)
library(dplyr)
library(scales)
test = read_tsv("~/tmp/20231122_downsample/dsample.tsv.gz",
    col_names=c("cg_num","seed","gsm","pattern","beta"))
test = test |> dplyr::filter(pattern %in% paste0("P", 1:2000)) |>
    dplyr::select(gsm, pattern, seed, cg_num, beta) |>
    pivot_wider(values_from="beta", names_from="pattern")
meta <- read_xlsx("~/cell_type/20220630_Liu2021_mouseBrainWGBS.xlsx")

patterns_to_test <- c(125,250,500,1000,2000)
reference_pattern <- system.file("extdata", "Liu2021_MouseBrain.cm", package = "MethScope")
input_pattern <- GenerateInput("/home/fuh1/intermediate/20240322_Liu2021_training.cg",reference_pattern)
sampled_celltype <- meta$MajorType[match(rownames(input_pattern), meta$SampleID)]
trained_model <- Input_training(input_pattern,sampled_celltype,number_patterns = 2000,cross_validation = F)
dfplot_merge <- c()
for (i in patterns_to_test){
  test_data = test[,paste0("P", 1:i)]
  new_cols <- as.data.frame(matrix(NA, nrow = nrow(test_data), ncol = (2000-i)))
  test_data <- cbind(test_data, new_cols)
  colnames(test_data) <- paste0("P", 1:2000)
  #trained_model <- Input_training(input_pattern,sampled_celltype,number_patterns = i,cross_validation = F)
  prediction_result <- PredictCellType(trained_model,test_data,smooth = F)
  test_meta = test[,1:3]
  test_meta$prediction = prediction_result$prediction_label
  test_meta$MajorType = meta$MajorType[match(test_meta$gsm, meta$SampleID)]
  dfplot = test_meta |> group_by(cg_num) |> summarize(n_match = sum(prediction == MajorType), n = n())
  dfplot$pattern_number <- i
  dfplot_merge <- rbind(dfplot_merge,dfplot)
}

#saveRDS(dfplot, file="~/cell_type/20250218_sparsity_pattern.rds")

#dfplot_merge <- readRDS("~/cell_type/20250218_sparsity_pattern.rds")
ggplot(dfplot_merge, aes(cg_num, n_match/n,color = as.factor(pattern_number))) + geom_line(aes(group = pattern_number), linewidth = 1) + geom_point() + scale_x_continuous(trans="log2",
    labels = trans_format("log2", math_format(2^.x))) + scale_y_continuous(labels=scales::percent,limits = c(0,1)) + xlab("#CpGs covered") + ylab("Overall Classification Accuracy") + theme_bw() + theme(text = element_text(size = 9),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "right")+labs(color = "Pattern Number")+ scale_colour_brewer(palette = "Set2")
ggsave("~/figures/fuh1/20250220_sparsitypatterns_sep.pdf",width=4,height=3)


test_data = test[,paste0("P", 1:2000)]
test_meta = test[,1:3]
test_meta$observed <- apply(test_data, 1, function(x) ncol(test_data)-sum(is.na(x)))

summary_df <- test_meta %>%
  group_by(cg_num) %>%
  summarize(
    mean_observed = mean(observed),
    sd_observed = 2*sd(observed),
    .groups = 'drop')

pdf("~/figures/fuh1/20250220_numberofpatterns.pdf",width=3,height=3,useDingbats = FALSE)
ggplot(summary_df, aes(x = cg_num, y = mean_observed)) +  geom_line(linewidth = 0.7,color="black") + geom_point(color = "black") + scale_x_continuous(trans="log2",
    labels = trans_format("log2", math_format(2^.x))) + geom_errorbar(aes(ymin = mean_observed - sd_observed, ymax = mean_observed + sd_observed),width = 0.1, color = "red") + xlab("#CpGs covered")+ theme_bw() + theme(text = element_text(size = 11),panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylab("#Patterns observed")
dev.off()
```


